Practical Machine Learning Project
========================================================

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

## Data 

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

## Resolution

To solve the problem, I'll first load the caret package, and as I have a 4 cores machine, I'll use the doSNOW library to use all of them. This will greatly reduce the computation time. So I'll declare the 4 cores cluster, and initialize it

```{r,warning=FALSE}
library(caret)
library(doSNOW)
cl<-makeCluster(4)
registerDoSNOW(cl)
```

Now I'll load the pml-training and pml-testing sets. The pml-testing will be used later for the submission part of the project.

```{r}
pml<-read.csv("pml-training.csv",stringsAsFactors=TRUE)
validation<-read.csv("pml-testing.csv",stringsAsFactors=TRUE)
```

A clean-up on the variable set is neccessary, so I eliminate all those which have any NA or empty string. Then, I'll also delete the variables at the start of the dataset that doesn't add any relevant information (rowid, timestamps, new window and  num_window). This process reduces the number of variables to 54
```{r}
keep<-function(x){!(any(is.na(x)) || any(x==""))}
pml<-pml[,sapply(pml,keep)]
pml<-pml[,-c(1,3,4,5,6,7)]
str(pml)
```

I'll divide the pml dataset into a training and testing datasets, on a 75-25 % proportion.
```{r}
set.seed(666)
inTrain <- createDataPartition(y=pml$classe,p=0.75, list=FALSE)
training <- pml[inTrain,]
testing <- pml[-inTrain,]
```

Then, I'll fit a model using random forest. This method doesn't need cross validation (https://www.stat.berkeley.edu/~breiman/RandomForests/cc_home.htm), so I'll define the train control with method 'none', and allowing parallel processing. The tuneGrid an tuneLength parameters are neccessary with method 'none'

```{r,warning=FALSE}
fit<-train(classe~., method="rf",data=training, trControl=trainControl(method="none",
        allowParallel = TRUE),tuneGrid = NULL, tuneLength = 1)
```

The model has a very high accuracy, with an 0.48% out of bag error rate estimate. 
```{r}
fit$finalModel
```

Plotting the model, it seems with ~100 trees instead of 500, it would have very similar performance 
```{r}
plot(fit$finalModel)
```

Now I'll check the out of sample error with the testing set.
```{r}
res<-predict(fit,newdata=testing)
confusionMatrix(res,testing$classe)
```
It shows a very high accuracy 99.71%, so I consider this model valid. I also tested with Boosting method (gbm) obtaining a 97% accuracy, as it was slightly worse, I finally decided for Random Forest. 

Then I'll stop the cluster with the 4 cores.
```{r}
stopCluster(cl)
```

And finally, I'll predict the 20 cases of the validation set, and send the results to the course.
```{r}
answers <-predict(fit,newdata=validation)
pml_write_files = function(x){
      n = length(x)
      for(i in 1:n){
            filename = paste0("problem_id_",i,".txt")
            write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
      }
}
pml_write_files(answers)
```

The model predicted correctly every one of the 20 cases.